Semantic Preserving Bijective Mappings for Representations of Special Functions between Computer Algebra Systems and Document Preparation Systems

### 2018-11-27 ReviewerResponse.txt

In the following we respond specifically to the reviewers comments.

Reviewer 1:

As suggested, we made several changes to improve the structure of the paper.  
In summary, we have :
(1) added subsection headings "Different Sets of Defined Functions", 
"Positions of Branch Cuts", "Insufficient Semantic Information", and 
"Ambiguous Expressions" in Section 3;
(2) gave an overview of the concept of the translator in Section 4;
(3) removed unnecessary Algorithm 1 in Section 5 as suggested, because 
it was just a simplified pre-version of Algorithm 2;
(4) added the subsection heading "Problems with the Look Ahead Approach" in 
Section 5 for a better outline; and 
(6) further explained the different states of Algorithm 3 (now Algorithm 2) 
by refer to the lines in the descriptions. We also explained how a power 
after a macro is handled in more detail and why optional parameters needs 
to be handled differently to other arguments.

@@

Note that we did not add further comparisons with LaTeXML to our 
evaluations section, because current CAS are unable to understand MathML 
which contains special functions. Thus we can only compare translations 
without function calls or implement a solution which overcomes this 
problem, in order to make a decent comparison. Because of the great amount 
of problems which would need to be overcome, a full solution translation 
implementation using MathML is out of the scope of our current project. 
This is now all explained in detail in the new version of the manuscript.

Reviewer 2:

For clarification, throughout the manuscript, we decided to change the 
wording "Word Processor" to "Document Preparation System." Since this wording
occurred as part of the title, we did not add that we focus on LaTeX to the 
title and hope that the Abstract points this out in an acceptable manner.

Also, instead of creating an Appendix, we decided to add short paragraphs 
which explain the purpose of the Sections 4 and 6. We hope 
this helps the reader to decide whether he can skip the section, e.g., in 
case of his lack of interest in Maple "hackings", or not.  In the new 
version of the paper, we implemented all suggested minor changes as well as 
adding the idea of improved numerical tests by using the mentioned cut 
region detection to the future work section.

### 2018-11-27 main.tex
Fix figure 4 wrapping (close #21)
Delted the phrase 'in consequence' all over the paper (close #22)

While \LaTeX{} focuses on displaying mathematics, a \gls*{cas} concentrates on computations and user friendly syntax. Especially important for a CAS is to embed unambiguous semantic information within the input. Each system uses different representations and syntax in consequence. Hence, a writer needs to constantly translate mathematical expressions from one representation to another and back again. Table~\ref{tab:JacobiP-usecase} shows four different representations for~(\ref{eq:P}).

@@

Hence, a \gls*{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls*{cas} to avoid mistakes during a translation~\parencite{Maple:Cuts}. In consequence, a manual translation process is not only laborious, but also prone to errors. Note that this general problem has been named as automatic \gls*{p2c} conversion~\parencite{POM-Tagger}.

@@

\section{Related Work}\label{sec:related-work}
Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls*{cas} provide simple functions to import and export mathematical \LaTeX{} expressions\footnote{The selected \gls*{cas} Maple, Mathematica, Matlab, and SageMath provide import and/or export functions for \LaTeX:\\
Maple, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;
Mathematica, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;
Matlab, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;
SageMath, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017.}. Those tools have two essential problems. They are only able to import simple mathematical expressions, where the semantics are unique. For example, the internal \LaTeX{} macro \verb|\frac| always indicates a fraction. However, for more complex expressions, e.g., the Jacobi polynomial in \Cref{tab:JacobiP-usecase}, the import functions fail. The second problem appears in the export tools. Mathematical expressions in \gls*{cas} are fully semantic. Otherwise the \gls*{cas} wouldn't be able to compute or evaluate the expressions. During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry sufficient semantic information. In consequence of these two problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above). Our tool should solve these problems and provide round-trip translations between \LaTeX{} and \gls*{cas}.

@@

\vspace{-0.2cm}
The position of the branch cut of the arccotangent function differs after translation~\eqref{eq:acot-alternatives}, which may lead to incorrect calculations later on. The alternative translations~\eqref{eq:acot-alternatives-1} and~\eqref{eq:acot-alternatives-2} use other functions instead of the arccotangent function. The arctangent function~\eqref{eq:acot-alternatives-1} and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts as in the \gls*{dlmf} and in \Maple. In consequence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.

@@

The parsed tree generated by the \gls*{pom}-Tagger is not a mathematical expression tree. The \gls*{pom} project aims to disambiguate mathematical \LaTeX{} expressions and generates an expression tree. However, in the current state, many expressions cannot be disambiguated yet. In consequence, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree. We call this parsed tree the \gls*{pom-pt}.

@@

Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to get a syntax tree of the input\footnote{In consequence, a license of \Maple{} is mandatory to perform backward translations. The translator is using the version \Maple{} 2016.}. \Maple{} allows several different input styles. The \texttt{1D} input is mainly used for programming purposes and therefore also used to perform our translations. Internally, \Maple{} uses a \gls*{dag} for syntax trees.

>>>>

While \LaTeX{} focuses on displaying mathematics, a \gls*{cas} concentrates on computations and user friendly syntax. Especially important for a CAS is to embed unambiguous semantic information within the input. Therefore, each system uses different representations and syntax, so that a writer needs to continually translate mathematical expressions from one representation to another and back again. Table~\ref{tab:JacobiP-usecase} shows four different representations for~(\ref{eq:P}).

@@

Hence, a \gls*{cas} user needs to fully understand the properties and special definitions (such as the position of branch cuts) in the \gls*{cas} to avoid mistakes during a translation~\parencite{Maple:Cuts}. Hence, a manual translation process is not only laborious, but also prone to errors. Note that this general problem has been named as automatic \gls*{p2c} conversion~\parencite{POM-Tagger}.

@@

\section{Related Work}\label{sec:related-work}
Since \LaTeX{} became the de facto standard for writing papers in mathematics, most of the \gls*{cas} provide simple functions to import and export mathematical \LaTeX{} expressions\footnote{The selected \gls*{cas} Maple, Mathematica, Matlab, and SageMath provide import and/or export functions for \LaTeX:\\
Maple, \url{http://www.maplesoft.com/support/help/Maple/view.aspx?path=latex} seen 06/2017;
Mathematica, \url{https://reference.wolfram.com/language/tutorial/GeneratingAndImportingTeX.html} seen 06/2017;
Matlab, \url{https://www.mathworks.com/help/symbolic/latex.html} seen 06/2017;
SageMath, \url{http://doc.sagemath.org/html/en/tutorial/latex.html} seen 06/2017.}. Those tools have two essential problems. They are only able to import simple mathematical expressions, where the semantics are unique. For example, the internal \LaTeX{} macro \verb|\frac| always indicates a fraction. However, for more complex expressions, e.g., the Jacobi polynomial in \Cref{tab:JacobiP-usecase}, the import functions fail. The second problem appears in the export tools. Mathematical expressions in \gls*{cas} are fully semantic. Otherwise the \gls*{cas} wouldn't be able to compute or evaluate the expressions. During the export process, the semantic information gets lost, because generic \LaTeX{} is not able to carry sufficient semantic information. Because of these two problems, an exported expression cannot be imported to the same system again in most cases (except for simple expressions such as those described above). Our tool should solve these problems and provide round-trip translations between \LaTeX{} and \gls*{cas}.

@@

\vspace{-0.2cm}
The position of the branch cut of the arccotangent function differs after translation~\eqref{eq:acot-alternatives}, which may lead to incorrect calculations later on. The alternative translations~\eqref{eq:acot-alternatives-1} and~\eqref{eq:acot-alternatives-2} use other functions instead of the arccotangent function. The arctangent function~\eqref{eq:acot-alternatives-1} and the natural logarithm~(\ref{eq:acot-alternatives-2}) have the same positioned branch cuts as in the \gls*{dlmf} and in \Maple. Hence, translation~(\ref{eq:acot-alternatives-1}) solves the issue, as long as the user does not evaulate the function at $z = 0$, while translation~(\ref{eq:acot-alternatives-2}) solves the issue except at $z = -\iunit$.

@@

The parsed tree generated by the \gls*{pom}-Tagger is not a mathematical expression tree. The \gls*{pom} project aims to disambiguate mathematical \LaTeX{} expressions and generates an expression tree. However, in the current state, many expressions cannot be disambiguated yet. Therefore, the \gls*{pom}-tagger generates a raw parsed tree where each token in the \LaTeX{} expression is a node in the tree. We call this parsed tree the \gls*{pom-pt}.

@@

Instead of writing a custom \Maple{} syntax parser, we use \Maple's internal data structure to get a syntax tree of the input\footnote{Therefore, a license of \Maple{} is mandatory to perform backward translations. The translator is using the version \Maple{} 2016.}. \Maple{} allows several different input styles. The \texttt{1D} input is mainly used for programming purposes and therefore also used to perform our translations. Internally, \Maple{} uses a \gls*{dag} for syntax trees.
